---
title: "Baseball"
author: "Marwa Mohamed"
date: "2/23/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# load needed packages and install them if they do not exist
easypackages::packages("tidyverse", prompt = F)
```

## R Markdown
Investigate how well we can predict the World Series winner at the beginning of the playoffs.

## Load data
```{r}
baseballTeams = read_csv("data/baseball.csv")
#dataset structure
str(baseballTeams)
#identify the total number of years included in this dataset.
table(baseballTeams$Year)

```

## Playoffs teams

```{r}
# a data frame limited to teams that made the playoffs 
baseball = subset(baseballTeams, Playoffs == 1)
#dataset structure
str(baseball)
#the number of teams making the playoffs in some season
PlayoffTable = table(baseball$Year)
PlayoffTable
#to get the names of the PlayoffTable vector
str(names(PlayoffTable))
#returns the number of playoff teams in 1990 and 2001
PlayoffTable[c("1990", "2001")]
#Store the number of teams in the playoffs for each team/year pair in the dataset
baseball$NumCompetitors = PlayoffTable[as.character(baseball$Year)]
str(baseball$NumCompetitors)
#the number of playoff team/year pairs are there in dataset from years where 8 teams were invited to the playoffs
baseball %>% 
  filter(NumCompetitors == 8) %>% 
  nrow()
```

## Bivariate Models for Predicting World Series Winner
```{r}
# whether a team won the World Series in the indicated year  (1 won, 0 otherwise)
baseball$WorldSeries = as.numeric(baseball$RankPlayoffs == 1)
#the number of teams that did NOT win the World Series
baseball %>% 
  filter(WorldSeries == 0) %>% 
  nrow()

```
Find variables that is a significant predictor of the WorldSeries variable in a bivariate logistic regression model( build 12 models)
```{r}
# year model
modyear = glm(WorldSeries ~ Year, data = baseball, family = binomial)
summary(modyear)
# RS model
modRS = glm(WorldSeries ~ RS, data = baseball, family = binomial)
summary(modRS)
# RA model
modRA = glm(WorldSeries ~ RA, data = baseball, family = binomial)
summary(modRA)
# W model
modW = glm(WorldSeries ~ W, data = baseball, family = binomial)
summary(modW)
# OBP model
modOBP = glm(WorldSeries ~ OBP, data = baseball, family = binomial)
summary(modOBP)
# SLG model
modSLG = glm(WorldSeries ~ SLG, data = baseball, family = binomial)
summary(modSLG)
# BA model
modBA = glm(WorldSeries ~ BA, data = baseball, family = binomial)
summary(modBA)
# RankSeason model
modRankSeason = glm(WorldSeries ~ RankSeason, data = baseball, family = binomial)
summary(modRankSeason)
# OOBP model
modOOBP = glm(WorldSeries ~ OOBP, data = baseball, family = binomial)
summary(modOOBP)
# OSLG model
modOSLG = glm(WorldSeries ~ OSLG, data = baseball, family = binomial)
summary(modOSLG)
# NumCompetitors model
modNumCompetitors = glm(WorldSeries ~ NumCompetitors, data = baseball, family = binomial)
#AIC 230.96
summary(modNumCompetitors)
# League model
modLeague = glm(WorldSeries ~ League, data = baseball, family = binomial)
summary(modLeague)

```
## Multivariate Models for Predicting World Series Winner
combine the variables found to be significant in bivariate models. Build a model using all of the variables that  found to be significant in the bivariate models
```{r}
mod = glm(WorldSeries ~ Year + RA + RankSeason + NumCompetitors, data = baseball, family = binomial)
summary(mod)
#correlation of significant vars
cor(baseball[,c(3,5,11,16)])

```

```{r}
#Year/RA model
mod1 = glm(WorldSeries ~ Year + RA, data = baseball, family = binomial)
summary(mod1)

#Year/RankSeason model
mod2 = glm(WorldSeries ~ Year + RankSeason, data = baseball, family = binomial)
summary(mod2)

#Year/NumCompetitors model
mod3 = glm(WorldSeries ~ Year + NumCompetitors, data = baseball, family = binomial)
summary(mod3)

#RA/RankSeason model
mod4 = glm(WorldSeries ~ RankSeason + RA, data = baseball, family = binomial)
summary(mod4)

#NumCompetitors/RA model
mod5 = glm(WorldSeries ~ NumCompetitors + RA, data = baseball, family = binomial)
summary(mod5)

#RankSeason/NumCompetitors model
mod6 = glm(WorldSeries ~ RankSeason + NumCompetitors, data = baseball, family = binomial)
summary(mod6)
```


