---
title: 'Data Wrangling Exercise 2: Dealing with missing values'
author: Mohammed Ali; Marwa Mohamed
date: 6/2/2019
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# load needed packages and install them if they do not exist
easypackages::packages("tidyverse", "stringi", "tidylog", prompt = F)
```

## Overview
In this exercise, you'll work with one of the most popular starter data sets in data science, the **Titanic** data set. This is a data set that records various attributes of passengers on the **Titanic**, including who survived and who didn’t.
Read the description of [the data set](https://www.kaggle.com/c/titanic/data) on the Kaggle website.

## Exercise

Using `R`, you’ll be handling missing values in this data set, and creating a new data set. Specifically, these are the tasks you need to do:

### 0: Load the data in RStudio

Save the data set `titanic3.xls` as a CSV file called `titanic_original.csv` and load it in RStudio into a data frame.

```{r}
titanic <- read_csv("data/titanic_original.csv", col_types = "iiffniifnfffif")

# Inspect data structure
glimpse(titanic)

# Inspect first 5 rows in loaded data
head(titanic)
```

### 1: Port of embarkation

The `embarked` column has some missing values, which are known to correspond to passengers who actually embarked at Southampton. Find the missing values and replace them with S. (Caution: Sometimes a missing value might be read into R as a blank or empty string.)

```{r}
# what is the status before processing
table(is.na(titanic$embarked))

# change missing values to s
titanic$embarked[is.na(titanic$embarked)] <- "S"

# after some cleaning
table(is.na(titanic$embarked))
```

### 2: Age

You’ll notice that a lot of the values in the `Age` column are either missing or illogic. While there are many ways to fill these missing values, using the *mean* or *median* of the rest of the values is quite common in such cases.

Calculate the mean of the Age column and use that value to populate the missing values

Think about other ways you could have populated the missing values in the age column. Why would you pick any of those over the mean (or not)?

```{r}
# age before cleaning
table(titanic$age)
# convert age to integers
titanic$age <- as.integer(titanic$age)
# replace missing and zero values with mean
titanic$age[is.na(titanic$age) | titanic$age == 0] <- mean(titanic$age, na.rm = TRUE)
# After cleaning
table(titanic$age)
```

### 3: Lifeboat

You’re interested in looking at the distribution of passengers in different `lifeboats`, but as we know, many passengers did not make it to a boat :-( This means that there are a lot of missing values in the `boat` column. Fill these empty slots with a dummy value e.g. the string *'None'* or *'NA'*

```{r}
# boat life before cleaning
table(is.na(titanic$boat))
# add none as factor to boat
levels(titanic$boat) <- c(levels(titanic$boat), 'None')
# replace missing boat life with NA
titanic$boat[is.na(titanic$boat)] <- 'None'
# After cleaning
table(is.na(titanic$boat))
```

### 4: Cabin

You notice that many passengers don’t have a cabin number associated with them.

- Does it make sense to fill missing cabin numbers with a value?

- What does a missing value here mean?

You have a hunch that the fact that the cabin number is missing might be a useful indicator of survival. Create a new column `has_cabin_number` which has *1* if there is a cabin number, and *0* otherwise.

```{r}
# Add new column with 0 values for missing cabin number and 1 otherwise
titanic$has_cabin_number <- ifelse(!(is.na(titanic$cabin)), 1, 0)
# Inspect data structure
glimpse(titanic)
```

