---
title: 'Data Wrangling Exercise 3: Human Activity Recognition'
author: "Mohammed Ali; Marwa Mohamed"
date: "6/2/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# load needed packages and install them if they do not exist
easypackages::packages("tidyverse", "tidylog", prompt = F)
```

## Overview
This data wrangling project is known to be a bit challenging since it's a good example of a messy, real-world data set that you'd encounter in a data science job.

The goal of this project is to get you some practice in processing real world datasets using the tools and techniques you have learnt so far.

You can download the dataset from [here](https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip). A full description of that data is available [here](http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones) and also in the README file included with the data.

This data set is organized in a way that makes it hard to use at first. You will need to use several data transformation techniques to put it into a usable, tidy state. Some extra guidelines and hints are in associated file called `1547028719_Samsung_data_wrangling_hints.pdf` in the data folder

## Exercise

### 0: Load the data in RStudio

Load the training and test data sets into RStudio, each in their own data frame.

```{r}
### Load train dataset
# Read subject_train file
subject  <- read_table("data/UCI HAR Dataset/train/subject_train.txt", col_names = FALSE)
# Add header
colnames(subject ) <- "subject"

#Read body_acc_x_train file
body_acc_x <- read_table("data/UCI HAR Dataset/train/Inertial Signals/body_acc_x_train.txt", col_names = FALSE)
# Create header eg.(body_acc_x_1)
header <- paste0(rep("body_acc_x_", 128), c(1:128))
# Add header
colnames(body_acc_x) <- header

#Read body_acc_y_train file
body_acc_y <- read_table("data/UCI HAR Dataset/train/Inertial Signals/body_acc_y_train.txt", col_names = FALSE)
# Create header eg.(body_acc_y_1)
header <- paste0(rep("body_acc_y_", 128), c(1:128))
# Add header
colnames(body_acc_y) <- header

#Read body_acc_z_train file
body_acc_z <- read_table("data/UCI HAR Dataset/train/Inertial Signals/body_acc_z_train.txt", col_names = FALSE)
# Create header eg.(body_acc_z_1)
header <- paste0(rep("body_acc_z_", 128), c(1:128))
# Add header
colnames(body_acc_z) <- header

#Read body_gyro_x_train file
body_gyro_x <- read_table("data/UCI HAR Dataset/train/Inertial Signals/body_gyro_x_train.txt", col_names = FALSE)
# Create header eg.(body_gyro_x_1)
header <- paste0(rep("body_gyro_x_", 128), c(1:128))
# Add header
colnames(body_gyro_x) <- header

#Read body_gyro_y_train file
body_gyro_y <- read_table("data/UCI HAR Dataset/train/Inertial Signals/body_gyro_y_train.txt", col_names = FALSE)
# Create header eg.(body_gyro_y_1)
header <- paste0(rep("body_gyro_y_", 128), c(1:128))
# Add header
colnames(body_gyro_y) <- header

#Read body_gyro_z_train file
body_gyro_z <- read_table("data/UCI HAR Dataset/train/Inertial Signals/body_gyro_z_train.txt", col_names = FALSE)
# Create header eg.(body_gyro_z_1)
header <- paste0(rep("body_gyro_z_", 128), c(1:128))
# Add header
colnames(body_gyro_z) <- header

#Read total_acc_x_train file
total_acc_x <- read_table("data/UCI HAR Dataset/train/Inertial Signals/total_acc_x_train.txt", col_names = FALSE)
# Create header eg.(total_acc_x_1)
header <- paste0(rep("total_acc_x_", 128), c(1:128))
# Add header
colnames(total_acc_x) <- header

#Read total_acc_y_train file
total_acc_y <- read_table("data/UCI HAR Dataset/train/Inertial Signals/total_acc_y_train.txt", col_names = FALSE)
# Create header eg.(total_acc_y_1)
header <- paste0(rep("total_acc_y_", 128), c(1:128))
# Add header
colnames(total_acc_y) <- header

#Read total_acc_z_train file
total_acc_z <- read_table("data/UCI HAR Dataset/train/Inertial Signals/total_acc_z_train.txt", col_names = FALSE)
# Create header eg.(total_acc_z_1)
header <- paste0(rep("total_acc_z_", 128), c(1:128))
# Add header
colnames(total_acc_z) <- header

#Read x train file
x  <- read_table("data/UCI HAR Dataset/train/X_train.txt", col_names = FALSE)
# Read features header
header <- t(read_table("data/UCI HAR Dataset/features.txt", col_names = FALSE))
# Add header
colnames(x) <- header

#Read y train file (activty numberss)
y <- read_table("data/UCI HAR Dataset/train/y_train.txt", col_names = FALSE)
# Create header
header <- "activity_number"
# Add header
colnames(y) <- header

# Create train data set
train <- data.frame(total_acc_x, total_acc_y, total_acc_z, body_acc_x, body_acc_y, body_acc_z, body_gyro_x, body_gyro_y, body_gyro_z, x, y, subject )
# Inspect train structure
glimpse(train)

### Load test dataset
# Read subject_test file
subject <- read_table("data/UCI HAR Dataset/test/subject_test.txt", col_names = FALSE)
# Add header
colnames(subject) <- "subject"

#Read body_acc_x_test file
body_acc_x <- read_table("data/UCI HAR Dataset/test/Inertial Signals/body_acc_x_test.txt", col_names = FALSE)
# Create header eg.(body_acc_x_1)
header <- paste0(rep("body_acc_x_", 128), c(1:128))
# Add header
colnames(body_acc_x) <- header

#Read body_acc_y_test file
body_acc_y <- read_table("data/UCI HAR Dataset/test/Inertial Signals/body_acc_y_test.txt", col_names = FALSE)
# Create header eg.(body_acc_y_1)
header <- paste0(rep("body_acc_y_", 128), c(1:128))
# Add header
colnames(body_acc_y) <- header

#Read body_acc_z_test file
body_acc_z <- read_table("data/UCI HAR Dataset/test/Inertial Signals/body_acc_z_test.txt", col_names = FALSE)
# Create header eg.(body_acc_z_1)
header <- paste0(rep("body_acc_z_", 128), c(1:128))
# Add header
colnames(body_acc_z) <- header

#Read body_gyro_x_test file
body_gyro_x <- read_table("data/UCI HAR Dataset/test/Inertial Signals/body_gyro_x_test.txt", col_names = FALSE)
# Create header eg.(body_gyro_x_1)
header <- paste0(rep("body_gyro_x_", 128), c(1:128))
# Add header
colnames(body_gyro_x) <- header

#Read body_gyro_y_test file
body_gyro_y <- read_table("data/UCI HAR Dataset/test/Inertial Signals/body_gyro_y_test.txt", col_names = FALSE)
# Create header eg.(body_gyro_y_1)
header <- paste0(rep("body_gyro_y_", 128), c(1:128))
# Add header
colnames(body_gyro_y) <- header

#Read body_gyro_z_test file
body_gyro_z <- read_table("data/UCI HAR Dataset/test/Inertial Signals/body_gyro_z_test.txt", col_names = FALSE)
# Create header eg.(body_gyro_z_1)
header <- paste0(rep("body_gyro_z_", 128), c(1:128))
# Add header
colnames(body_gyro_z) <- header

#Read total_acc_x_test file
total_acc_x <- read_table("data/UCI HAR Dataset/test/Inertial Signals/total_acc_x_test.txt", col_names = FALSE)
# Create header eg.(total_acc_x_1)
header <- paste0(rep("total_acc_x_", 128), c(1:128))
# Add header
colnames(total_acc_x) <- header

#Read total_acc_y_test file
total_acc_y <- read_table("data/UCI HAR Dataset/test/Inertial Signals/total_acc_y_test.txt", col_names = FALSE)
# Create header eg.(total_acc_y_1)
header <- paste0(rep("total_acc_y_", 128), c(1:128))
# Add header
colnames(total_acc_y) <- header

#Read total_acc_z_test file
total_acc_z <- read_table("data/UCI HAR Dataset/test/Inertial Signals/total_acc_z_test.txt", col_names = FALSE)
# Create header eg.(total_acc_z_1)
header <- paste0(rep("total_acc_z_", 128), c(1:128))
# Add header
colnames(total_acc_z) <- header

#Read x test file
x <- read_table("data/UCI HAR Dataset/test/X_test.txt", col_names = FALSE)
# Read features header
header <- t(read_table("data/UCI HAR Dataset/features.txt", col_names = FALSE))
# Add header
colnames(x) <- header

#Read y test file (activty numbers)
y <- read_table("data/UCI HAR Dataset/test/y_test.txt", col_names = FALSE)
# Create header
header <- "activity_number"
# Add header
colnames(y) <- header

# Create test data set
test <- data.frame(total_acc_x, total_acc_y, total_acc_z, body_acc_x, body_acc_y, body_acc_z, body_gyro_x, body_gyro_y, body_gyro_z, x, y, subject)
# Inspect test structure
glimpse(test)


```


### 1: Merge data sets

Merge the training and the test sets to create one data set.

```{r}
# Combine both sets
comb <- bind_rows(train, test)
#Inspect the first 5 rows in loaded data
head(comb)
```

2: Mean and standard deviation

Create two new columns, containing the mean and standard deviation for each measurement respectively. *Hint: Since some feature/column names are repeated, you may need to use the make.names() function in R*.

```{r}
#make.names(colnames(comb), unique = TRUE)
#a <- comb %>%
#  group_by(activity_number) %>%
#  select(contains("mean")) %>%
#  mutate(mean = list(colMeans(., na.rm = TRUE)))

```

### 3: Add new variables

Create variables called `ActivityLabel` and `ActivityName` that label all observations with the corresponding activity labels and names respectively

```{r}
# Add ActivityLabel column
comb <- comb %>% 
  mutate(ActivityLabel = case_when(activity_number == "1" ~ "Walking",                                                                        activity_number == "2" ~ "Walking_Upstairs",
                                   activity_number == "3" ~ "Walking_ownstairs",
                                   activity_number == "4" ~ "Sitting",
                                   activity_number == "5" ~ "Standing",
                                   activity_number == "6" ~ "Laying"),
         ActivityLabel = factor(ActivityLabel, levels = c("Walking", "Walking_Upstairs", "Walkind_Downstairs", "Sitting", "Standing", "Laying")))
# Inspect data structure
glimpse(comb)
```

### 4: Create tidy data set

From the data set in step 3, creates a second, independent tidy data set with the average of each variable for each activity and each subject. 

```{r}
```

