---
title: "Music_Records"
author: "Marwa Mohamed"
date: "2/18/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# load needed packages and install them if they do not exist
easypackages::packages("tidyverse", prompt = F)
```

## R Markdown

The music industry has a well-developed market with a global annual revenue around $15 billion. The recording industry is highly competitive and is dominated by three big production companies which make up nearly 82% of the total annual album sales.

Artists are at the core of the music industry and record labels provide them with the necessary resources to sell their music on a large scale. A record label incurs numerous costs (studio recording, marketing, distribution, and touring) in exchange for a percentage of the profits from album sales, singles and concert tickets.

Unfortunately, the success of an artist's release is highly uncertain: a single may be extremely popular, resulting in widespread radio play and digital downloads, while another single may turn out quite unpopular, and therefore unprofitable.

Knowing the competitive nature of the recording industry, record labels face the fundamental decision problem of which musical releases to support to maximize their financial success.

How can we use analytics to predict the popularity of a song? In this assignment, we challenge ourselves to predict whether a song will reach a spot in the Top 10 of the Billboard Hot 100 Chart.

Taking an analytics approach, we aim to use information about a song's properties to predict its popularity. The dataset songs (CSV) consists of all songs which made it to the Top 10 of the Billboard Hot 100 Chart from 1990-2010 plus a sample of additional songs that didn't make the Top 10.

### 0: Load the data in RStudio
```{r}
songs = read_csv("data/songs.csv")
#dataset structure
str(songs)
```

### Understanding the Data
How many songs does the dataset include for which the artist name is "Michael Jackson"?
```{r}
count(songs, artistname == "Michael Jackson")
```
### divide data

```{r}
SongsTrain = filter(songs, year <= 2009)
str(SongsTrain)
SongsTest = filter(songs, year == 2010)
str(SongsTest)
```

### Creating Prediction Model
 exclude some of the variables in dataset from being used as independent variables 
```{r}
nonvars = c("year", "songtitle", "artistname", "songID", "artistID")

# To remove these variables from your training and testing sets, type the following commands in your R console:

SongsTrain = SongsTrain[ , !(names(SongsTrain) %in% nonvars) ]

SongsTest = SongsTest[ , !(names(SongsTest) %in% nonvars) ]

```

build a logistic regression model
```{r}
SongsLog1 = glm(Top10 ~ ., data=SongsTrain, family=binomial)
summary(SongsLog1)
#Conclusion
#20 variables have at least one star next to the p-values, which represents significance at the 5% level.
#The higher our confidence about time signature, key and tempo, the more likely the song is to be in the Top 10
#Mainstream listeners tend to prefer less complex songs
#Mainstream listeners prefer songs with heavy instrumentation

```
### Validating Model
```{r}
SongsTestPred = predict(SongsLog1, newdata = SongsTest, type = "response")
#create a confusion matrix with a threshold of 0.45 
table(SongsTest$Top10, SongsTestPred >= 0.45)
#accuracy
(309+15)/(309+5+44+15)
# accuracy of the baseline model be on the test set(an easier model would be to pick the most frequent outcome (a song is not a Top 10 hit) for all songs.) 
table(SongsTest$Top10)
#accuracy
314/(314+59)
#True Positive Rate of model on the test set, using a threshold of 0.45
15/(15+44)
# False Positive Rate of model on the test set, using a threshold of 0.45?
5/(309+5)
```

